---
title: "RDD"
author: "Filip Mellgren"
date: '2019-11-01'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, include=FALSE}
library(rio)
library(tidyverse)
library(locfit)
library(rddensity)
library(rdd)
library(stargazer)
```

```{r airbnb}
df <- import("chic_aug_07.csv")
df %>% mutate(PriceX = gsub("[^0-9.]", "", df$price_string),
              PriceX = as.numeric(PriceX),
              dist_t = avg_rating - round(avg_rating/0.5)*0.5,
              treat = if_else(dist_t == 0, 1, 0)) -> df
# Rating, log price relationship
df %>% filter(avg_rating > 0) %>% ggplot(aes(x = avg_rating, y = log(PriceX))) + 
  geom_point(alpha = 0.3, color = "turquoise") + geom_smooth(method = "lm", formula = y~x) +
  # transpaency
  theme(
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
        )
ggsave("correlation.png", bg = "transparent")

# "Discontinuity plot"
df %>% ggplot(aes(x = dist_t, y = log(PriceX))) + 
  geom_point(alpha = 0.3, color = "turquoise") +
  stat_summary(fun.y = "mean", geom = "point", color = "coral", size = 2) +
  labs(title = " No discontiuity around threshold", 
       x = "Distance to rounding cut off", y = "Log nightly price"
       )+
  # transpaency
  theme(
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
        )
ggsave("airbnb_rd.png", bg = "transparent")


```



# New analysis, replicating Manacorda, Miguel, Vigorito (2009)

Paper can be accessed here:
http://emiguel.econ.berkeley.edu/assets/miguel_research/17/_WorkingPaper__Government__Transfers__and__Political_Support.pdf

Abstract of the paper:
"We estimate the impact of a large anti-poverty cash transfer program, the Uruguayan PANES, on political
support for the government that implemented it. Using the discontinuity in program assignment based
on a pre-treatment eligibility score, we find that beneficiary households are 11 to 14 percentage points
more likely to favor the current government relative to the previous government. Political support
effects persist after the program ends. A calibration exercise indicates that these persistent impacts
are consistent with a model of rational but poorly informed voters learning about politicians’ redistributive
preferences."

It strives to establish causal effects of government benefits on government support. The following is noted regarding the current academic debate about the issue:

* There is an empirically established relationship that voters respond to macro economic trends.
* At the same time, there are econometric concerns about endogeneity and small, aggregated samples.
* The true, causal, magnitude is unknown.

In their paper, the authors wishes to establish a causal link between political support of the incumbent party and government transfers. The setting is such that families receive a government program called "PANES" if their predicted income is above a certain threshold. The government bases its support on a prediction because reported income may suffer from being unstable following a major crisis in the country in the foremath of the event, hard to verify claims as the target often worked informally, and multiple variables help mitigate strategic misreporting.

In summary the Panes program looked had the following defining features:

* It was a temporary cash transfer and food voucher following an economic downturn in the year 2002
* Recipients were households with low predicted income "based on a large number of pre-treatment covariates".
- following crisis, many had an unstable income, so present was bad precitor of permanent income
- hard to verify income claims as the target populaiton of worked informaly
- multiple variables helps against strategic misreporting
* Households around the predicted income threshold for receiving Panes were "surveyed and asked a series of questions including their support for the current government, and a second similar follow-up survey took place the following year".


```{r load_data}
#df <- import("vn_estimation_2016.dta")
#df_pres <- import("Government Transfers_Replication/conf_president.dta")
df_mccrary <- import("Government Transfers_Replication/mccrary.dta")
df_reg_panes <- import("Government Transfers_Replication/reg_panes.dta")
```

```{r data_manipulation}
# Create variables indicating whether above or below threshold
df_reg_panes %>% mutate(above = if_else(ind_reest > 0, ind_reest, NaN),
                        below = if_else(ind_reest < 0, ind_reest, NaN),
                        gov2007 = case_when(
                          h_89 == 9 ~NaN,
                          h_89 == NA ~NaN,
                          TRUE ~ (h_89-1)/2)
                        ) -> df
```



## Descriptive Statistics

To begin the analysis, I establish some observational facts by estimating a regression using OLS. Of course, these results will contain endogeneity bias but will nonetheless be useful as a benchmark. Potential reasons for endogeneity might include:
* first
* second
* third

```{r exploration}
# Simple regression, approval status on predicted income
ols_fit <- lm(aprobado ~ ind_reest, data = df)
stargazer(ols_fit, type = "html")
```

Overall, there is a negative relationship between predicted income and government support. Likely reflecting party difference in support among various income groups.

## Exogenous variation
To escape from the endogeneity present in the previous analysis, we can think of predicted income as a variable with some exogenous variation around the threshold of being assigned the government support. The basic idea is that households on either side of the threshold will be similar in all characteristics but their treatment status. 

Falling slightly to the right means some indicators of income predicts the household is above the threshold whereas falling on the left side of the threshold means the predcitors state that the household is poor enough to receive the benefits. However, except for extremely modest variation in the income indicators, predicted income may be deemed as something varying almost entirely randomly. 

If there happens to be a large discontinuity in government support around the specified threshold, the effect can plausibly be attributed to treatment status as nothing else may be expected to vary around the somewhat arbtirary threshold of predicted income.

There are, however, a few issues that must be dealt with. First, if certain households anticipate that they may be close to the threshold, they might try to affect the measure predicted income so as to move themselves to the left and thus increase their chance of receiving the governemtn benefit resulting in a potnetially biased sample where the treated population is skewed towards households capable of affecting the outcomes of the government transfer, A second issue relate to potential confounding variables that also vary around the threshold. For example, if it turns out that predicted income was also used for some other government program, then it'd be unclear from what program potnetial effects adhere. Another issue might be if the metric varies around the threshold based on some observable characteristics, then effects could depend on this characteristic rather than treatment status. 

These issues, however, can be checked. The first issue implies a skewed density distirbution around the threshold whereas the second needs to make the case that this variable exists and then show a discontinuity precisly around the threshold. In practice, it is hard to come up with such variables and especially in this case because predicted income is a construct of several variables and unlikely to be used elsewhere.

A final concern partains to whether there is a discontinuity or just a sharp (continuous) decline. This will be an issue if the so called "bandwidth" is wide (See image). 

There is a connection between a regression discontinuity estimates and instrumental variables (in fact, an RDD is an IV estimation). I therefore proceed by presenting the analysis as if it was an IV together with the rich set of intuitive graphs present in an RDD estimation.

Add in: "A concern is manipulation of program assignment by either officials
or enumerators, due to strategic responses or a correlation between survey non-response and
political views."

Another concern: did recipients report false non truthful values about their government support?

## First stage
The first stage in an IV study considers the strength of the exogenous instrument $z$ (referred to as the "instrument"") in predicting the endogenous variable of interest $x$. Ideally, the instrument is a strong predictor of the endogenous variable to avoid issues related to weak variable bias (Bound Jaeger Baker (1995)). Other than establishing a tight link between the two variables, the first stage is simply a mechanic check that needs to be done. It's importance is measured by an F statistic which is ideally larger than 10.

```{r}
# First stage
df %>% gather(above, below, key = "Treatment", value = "Predicted_income") %>% 
  ggplot(aes(x = Predicted_income, y = aprobado, 
             color = Treatment, fill = Treatment)) +
  geom_smooth(method = "loess", span = 0.5) +
  geom_point(size = 0.3) +
  theme_minimal() +
  labs(title = " Discontiuity around threshold", 
       x = "Predicted income", y = "Ever received PANES") +
    theme(
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
        )
ggsave("first_stage.png", bg = "transparent")
```
As can be seen in the figure above, there is a clear relationship between the two. Namely that when predicted income changes status, so does PANES status. This is intuitively clear because the government is only handing out the PANES to eligible households (those poor according to predicted income). There are a few exceptions but overall, the relationship is very strong which in the words of RDD can be referred to as a "Sharp RDD".

## Reduced form, support for government

```{r}
# Reduced form: support for govt - first wave
# Lower span is less biased, but a noisier estimate.

df %>% 
  gather(above, below, key = "Treatment", value = "Predicted_income") %>% 
  ggplot(aes(x = Predicted_income, y = gov2007, 
             color = Treatment, fill = Treatment)) +
  geom_smooth(method = lm, span = 0.67) +
  theme_minimal() +
  labs(title = "Linear estimate", 
       x = "Predicted income", y = "Government support 2007") +
    theme(
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
        )
ggsave("reduced_line.png", bg = "transparent")


df %>% 
  gather(above, below, key = "Treatment", value = "Predicted_income") %>% 
  ggplot(aes(x = Predicted_income, y = gov2007, 
             color = Treatment, fill = Treatment)) +
  geom_smooth(method = "loess", span = 0.33) +
  theme_classic() +
  labs(title = "Span = 0.33", 
       x = "Predicted income", y = "Government support 2007") +
    theme(
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
        )
ggsave("reduced_0.33.png", bg = "transparent")

df %>% 
  gather(above, below, key = "Treatment", value = "Predicted_income") %>% 
  ggplot(aes(x = Predicted_income, y = gov2007, 
             color = Treatment, fill = Treatment)) +
  geom_smooth(method = "loess", span = 0.67) +
  theme_minimal() +
  labs(title = "Span = 0.67", 
       x = "Predicted income", y = "Government support 2007") +
    theme(
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
        )
ggsave("reduced_0.67.png", bg = "transparent")
  
```


## Other variables varying around the threshold



## Density around threshold
```{r}
df %>% 
  ggplot(aes(x = ind_reest)) + 
  geom_density(trim = TRUE, alpha = 0.4) +
  geom_histogram(aes(y = ..density..), bins = 30, alpha = 0.5)
```
```{r}
rdd <- rddensity(X = df$ind_reest)

# Plot
plot <- rdplotdensity(rdd, df$ind_reest, plotRange = c(-0.02, 0.02), plotN = 25)
  

```
```{r}

df_mccrary %>% mutate(tot = sum(ind_reest),
                      pct = quantile(ind_reest, c(.99)),
                      ind_reest = if_else(ind_reest < pct, ind_reest, NaN)
                      ) -> df_mccrary
  

Mcrary_test <- DCdensity(df_mccrary$ind_reest, 0, bin = NULL, bw = NULL, verbose = FALSE, plot = TRUE, ext.out = FALSE, htest = FALSE)


```



## Bandwidth

# Conclusion



